#Practical Machine Learning - Course Project
### Ashwin Kadle
### 8th January 2017

### Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

### Data Source

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment. 

### Downloading and Loading the Dataset

The pml-training.csv data is actually used to devise training and testing sets. The pml-test.csv data is used to predict and answer the 20 questions based on the trained model.

```{r}
training.file   <- 'pml-training.csv'
test.cases.file <- 'pml-testing.csv'
training.url    <- 'http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv'
test.cases.url  <- 'http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv'

download.file(training.url, training.file)
download.file(test.cases.url,test.cases.file )

```
### Cleaning Data
After Initial Exploratory Analysis the dataset has been modified to convert Blanks,#DIV/0 and NA values to NA

```{r}
training   <-read.csv(training.file, na.strings=c("NA","#DIV/0!", ""))
testing <-read.csv(test.cases.file , na.strings=c("NA", "#DIV/0!", ""))
training<-training[,colSums(is.na(training)) == 0]
testing <-testing[,colSums(is.na(testing)) == 0]

```

The fields not related to calculation are removed from the dataset

```{r}
training<-training[,-c(1:7)]
testing<-testing[,-c(1:7)]
```

### Random sampling of data into training and test sets

```{r}
set.seed(1)
library("caret")
inTrain <- createDataPartition(y=training$classe, p=0.75, list=FALSE)
training.data <- training[inTrain, ]
testing.data  <- training[-inTrain, ]
```

### Data Manipulation Building Decision Tree Model

```{r}
library("rpart")
library("rattle")
modFitDT <- rpart(classe ~ ., data = training.data, method="class")
fancyRpartPlot(modFitDT)
```

### Predicting with Decision Tree
```{r}
prediction <- predict(modFitDT, testing.data, type = "class")
confusionMatrix(prediction, testing.data$classe)
```

The Accuracy is 0.74

###  Building the Random Forest Model

```{r}
library("randomForest")
modFitRF <- randomForest(classe ~ ., data = training.data, ntree = 1000)
prediction <- predict(modFitRF, testing.data, type = "class")
confusionMatrix(prediction, testing.data$classe)
```

The accuracy is 0.99

### Out of sample error rate

```{r}
missClass = function(values, predicted) {
  sum(predicted != values) / length(values)
}
OOS_errRate = missClass(testing.data$classe, prediction)
OOS_errRate
```

### Predicting on Testing Data Predicting with Decision Tree

```{r}
predictionDT <- predict(modFitDT,testing, type = "class")
predictionDT
```

### Predicting with Random Forest

```{r}
predictionRF <- predict(modFitRF, testing, type = "class")
predictionRF
```

### Submission
```{r}
write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_",i,".txt")
    write.table(x[i], file=filename, quote=FALSE, row.names=FALSE, col.names=FALSE)
  }
}
write_files(predictionRF)
```









